---
title: "Class 5: Data Visualization with ggplot"
author: "Aida (PID: A59026788)"
format: html
---

## Using GGPLOT

To use ggplot2 we first need to install it on our computers. To do this we will use the function `install.packages()`

Before I use any package functions I have to load them up with a `library()` call (refer to below)

```{r}
library(ggplot2)
ggplot(cars)
```

```{r}
cars
```

There is always the "base R" graphics system, i. e. `plot()`

```{r}
plot(cars)
```

To use ggplot I need to spell out at least 3 things:
- data (the stuff I want to plot as a data.frame)
- aesthetics (aes() values - how the data map to the plot)
- geoms (how I want things drawn)

```{r}
ggplot(cars) + 
  aes(x=speed, y=dist) + 
  geom_point() +
  labs(title = "Speed and Stopping Distances of Cars", 
       x="Speed (MPH)", 
       y="Stopping Distance (ft)", 
       subtitle = "Your informative subtitle text here", 
       caption = "dataset:`cars`")
  geom_smooth(method="lm", se=FALSE)
  theme_bw()
```

```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```

```{r}
nrow(genes)
```


```{r}
colnames(genes)
ncol(genes)
```


```{r}
table(genes$State)
```


```{r}
round(table(genes$State)/nrow(genes) * 100, 2)
```


```{r}
ggplot(genes) +
  aes(x=Condition1, y=Condition2) +
  geom_point()
```

```{r}
ggplot(genes) +
  aes(x=Condition1, y=Condition2, col=State) +
  geom_point()
```


```{r}
p <- ggplot(genes) +
  aes(x=Condition1, y=Condition2, col=State) +
  geom_point()
```




```{r}
p + scale_colour_manual(values = c("blue", "gray", "red"))
```

```{r}
p + scale_colour_manual(values = c("blue", "gray", "red")) + 
  labs(title="Gene expression Changes Upon Drug Treatment", x="Control (no drug)", y="Drug Treatment")
```

```{r}
# File location online
url <- "https://raw.githubusercontent.com/jennybc/gapminder/master/inst/extdata/gapminder.tsv"

gapminder <- read.delim(url)
```


```{r}
library(dplyr)
```

```{r}
gapminder_2007 <- gapminder %>% filter(year==2007)
```

```{r}
ggplot(gapminder_2007) +
  aes(x=gdpPercap, y=lifeExp) +
  geom_point()
```

```{r}
ggplot(gapminder_2007) +
  aes(x=gdpPercap, y=lifeExp) +
  geom_point(alpha=0.5)
```


```{r}
ggplot(gapminder_2007) +
  aes(x=gdpPercap, y=lifeExp, color=continent, size=pop) +
  geom_point(alpha=0.5)
```


```{r}
ggplot(gapminder_2007) +
  aes(x=gdpPercap, y=lifeExp, color=pop) +
  geom_point(alpha=0.5)
```


```{r}
ggplot(gapminder_2007) +
  aes(x=gdpPercap, y=lifeExp, size=pop) +
  geom_point(alpha=0.5)
```


```{r}
ggplot(gapminder_2007) +
  geom_point(aes(x=gdpPercap, y=lifeExp, size=pop, alpha=0.5)) +
  scale_size_area(max_size = 10)
```


```{r}
gapminder_1957 <- gapminder %>% filter(year==1957)

ggplot(gapminder_1957) +
  aes(x=gdpPercap, y=lifeExp, color=continent, size=pop) +
  geom_point(alpha=0.7) +
  scale_size_area(max_size = 10)
```


```{r}
gapminder_1957 <- gapminder %>% filter(year==1957 | year==2007)

ggplot(gapminder_1957) +
  geom_point(aes(x=gdpPercap, y=lifeExp, color=continent, size=pop, alpha=0.7)) +
  scale_size_area(max_size = 10) +
  facet_wrap(~year)
```



```{r}
ggplot(mtcars) + aes(x=mpg, y=disp) + geom_point()
```


```{r}
ggplot(mtcars, aes(mpg, disp)) + geom_point()
```

